MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads

### 2017-03-07 mlppi-comsis17/paper.tex
MLPPI updates

(Another possibility would be to use the partitioning associated with {\tt LO\_DISCOUNT} shown 
in the beginning of this section with only two partitions.)

@@

%\subsection{Solution: An Automated Partitioning Tool}
%\label{sec:solution}
\paragraph{Solution:} 
To ease a burden of partitioning decision, 
we introduce a novel physical database design tool, called {\em MLPPI wizard}, 
designed to recommend an effective partitioning solution for a given workload.
%Indeed, the \hbox{wizard} is the first physical database design tool 
%developed at Teradata.
This tool uses a greedy algorithm for search space enumeration 
and is based on a general framework allowing general expressions, 
ranges and case expressions for partition definitions. 
%is particularly well-suited for an MLPPI definition. 
We emphasize that the \hbox{predicate-driven} algorithm used by the tool can be applied 
to {\em any} clustering or partitioning scheme based on simple fields and expressions or complex SQL predicates 
in an {\em arbitrary} DBMS. 
The wizard also borrows the optimizer's cost model 
to prune the search space and reach a final solution.

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.48]{architecture}
\vspace{-0.1in}
\caption{The MLPPI wizard architecture\label{fig:arch}}
\end{figure*}

Figure~\ref{fig:arch} illustrates how our tool recommends 
the final MLPPI customized for a given workload consisting of 
a set of queries and corresponding weights. 
The wizard passes through a series of phases: {\em Preprocessing}, 
{\em Initial}, and {\em Optimized} Phases. 
In the pre-processing phase the tool produces the most granular partitioning (MGP) based on the predicates collected from every query.
In the initial phase the wizard refines the passed MGP 
by merging a pair of partitions with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}) 
and then yields an initial (feasible) MLPPI based on the refined partitions.
In the \hbox{optimization} phase, the tool makes a further attempt to optimize the initial MLPPI. 
We elaborate in greater detail on each phase, 
using the running query set $Q$ consisting of $q$1 and $q2$ in the rest of the article.

%\shorten{
%An initial MLPPI is produced through the top two phases reflecting practical constraints by field or partition. 
%It is further optimized in the last phase.
%In the preprocessing phase, 
%for each query its predicate(s) is extracted via an API call (on {\tt explain}) made to the server. 
%Then, the wizard produces the most granular partitioning (MGP) based on the 
%predicates collected from every query. 
%If the MGP has too many fields over a field count limit, 
%then the wizard \hbox{incrementally} throws away a field from the MGP until the limit. 
%
%In the initial phase, the passed MGP gets reduced if it includes more partitions than a partition count limit allowed by an MLPPI. 
%The refinement can be \hbox{incrementally} done by merging a pair of partitions 
%with the least {\em scan cost} (as will be explained in \hbox{Section~\ref{sec:init_phase}}), 
%until the limit. 
%After the phase, an initial (feasible) MLPPI is produced based on the refined partitions.
%
%In the optimization phase, the tool makes a further attempt to optimize the initial MLPPI. 
%Similar to the initial phase, more partitions can be reduced by merging ones with the 
%least {\em query cost} (as will be described in Section~\ref{sec:opt_phase}). 
%If the ongoing solution cannot be further improved, the tool makes the final MLPPI recommendation. 
%We will provide more details about each phase later, 
%using the query set $Q$ consisting of $q$1 and $q2$ 
%in the rest of the article.}

To the best of our knowledge, there is few extant industrial tool 
to addressing the {\it {\bf \hbox{multi-level} partitioning}} problem, except our wizard. 
The wizard resides completely on the client side as opposed to 
several previous tools~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} 
that require optimizer code instrumentation. 
Our tool simply uses existing APIs to simplify the queries, capture fact table 
predicates and costs, and uses these items to make a recommendation. 
%This makes the wizard extensible and portable to different releases 
%of the Teradata DBMS server. 
This \hbox{loosely-coupled} approach makes the wizard extensible and portable to different releases 
of the database server.

>>>>

%(Another possibility would be to use the partitioning associated with {\tt LO\_DISCOUNT} shown 
%in the beginning of this section with only two partitions.)

@@

\paragraph{Solution:} 
To ease a burden of partitioning decision, 
we introduce a novel physical database design tool, called {\em MLPPI wizard}, 
designed to recommend an effective partitioning solution for a given workload.
%Indeed, the \hbox{wizard} is the first physical database design tool 
%developed at Teradata.
This tool uses a greedy algorithm for search space enumeration 
and is based on a general framework allowing general expressions, 
ranges and case expressions for partition definitions. 
%is particularly well-suited for an MLPPI definition. 
We emphasize that the \hbox{predicate-driven} algorithm used by the tool can be applied 
to {\em any} clustering or partitioning scheme based on simple fields and expressions or complex SQL predicates 
in an {\em arbitrary} DBMS. 
The wizard also borrows the optimizer's cost model 
to prune the search space and reach a final solution.

\begin{figure*}[htp!]
\centering
\includegraphics[scale=0.45]{architecture}
\vspace{-0.1in}
\caption{The MLPPI wizard architecture\label{fig:arch}}
\end{figure*}

Figure~\ref{fig:arch} illustrates how our tool recommends 
the final MLPPI customized for a given workload consisting of 
a set of queries and corresponding weights. 
Until yielding a final partitioning recommendation, 
the wizard passes through a series of phases: {\em Preprocessing}, 
{\em Initial}, and {\em Optimized} Phases. 
%In the pre-processing phase the tool produces the most granular partitioning (MGP) based on the predicates collected from every query.
%In the initial phase the wizard refines the passed MGP 
%by merging a pair of partitions with the least (scan) cost 
%and then yields an initial (feasible) MLPPI based on the refined partitions.
%In the \hbox{optimization} phase, the tool makes a further attempt to optimize the initial MLPPI. 
We elaborate in greater detail on each of the three phases, 
using the running query set $Q$ consisting of $q$1 and $q2$ in the rest of the article. 
As illustrated in Figure~\ref{fig:arch}, the wizard resides completely on the client side as opposed to 
several previous tools~\cite{agrawal04:integrating,Lightstone04:db2auto,nehme2011automated} 
that require optimizer code instrumentation. 
Our tool simply invokes the server's existing APIs used to simplify the queries, capture fact table 
predicates, and estimate processing costs. 
%This makes the wizard extensible and portable to different releases 
%of the Teradata DBMS server. 
%This \hbox{loosely-coupled} interaction 
%between both sides makes the wizard extensible and portable to different releases 
%of the database server. 
To the best of our knowledge, our wizard  
is the very first physical database design tool 
to address such a {\em \hbox{multi-level} partitioning} problem.

\paragraph{Contribution:} This article provides 
the following contributions.

\begin{itemize}
\item We propose a physical database design tool, termed MLPPI wizard, to recommend a multi-level partitioning definition 
for reducing the processing cost of a given query workload on a star schema benchmark. 
\item This tool is totally outside the database server, thereby incurring no overhead of instrumenting query optimizer's code.
\item We empirically show that a partitioning recommendation of our tool outperforms that of human expert 
over increasing workload size and growing fact table.

\end{itemize}
